Compositional equivalences based on Open pNets

### 2020-02-24 2016-Weak-FH-Bisim/WeakBisim.tex
work on appendix

%% \TODO{voisr si on redefinit les bigotimes pour prendre des maps et retourner des maps au lieu des substitutions (et on englobe dans une subst a la fin)}

%% \TODO{prendre soin des posts sur les actions partout}

%% ===> dealine 14 Juillet

%% TODOs du 19 decembre

%% - enlever 5.2 et citer avocs

%% - ajouter des proof sketch dans le papier

%% - def 11 point 2:
%% A, Pred, Post
%% -----------   in T
%% s->s'
%% et A(j)=tau
%% implies blablabla

%% OLD OLD OLD (check)
%% Notes de reunion 24 Sept

%% - etat de l art: il faudrait des refs sur bisim en gemneral, probablement avec pas mal de pi calcul (Eric?), des refs sur pourquoi pnets et des refs sur utilite de la weak bisim en verif

%% - target: JLamp

%% Conventions
%% [14/9/2018]

%% Dans le texte pas de $\backslash$ devant pNet et pLTS

%% pNet devient P ou Q dans les equations; on garde $pLTS$

%% \medskip

%% TODOS et deadline

%% Global and questions:
%% \begin{itemize}
%% \item \TODO{should we use iff/implies or LeftRightArrow/Rightarrow}
%% \item related works
%% \end{itemize}
%% Ludo:
%% \begin{itemize}
%% \item[X] reorga sec weak bisim + thm -> 10 Juil
%% \item[X]  intro et orga globale du papier -> fin aout
%% \item[partial] explication dans semantique open pnets et composition
%% \item uniformisation pNet P Q
%% \end{itemize}

@@

Fig. \ref{SimpleProt:Spec} (left) shows the example principle: two unspecified
processes $P$ and $Q$ (holes) communicate messages, with a data value
argument, through the two protocol entities. Process $P$ send an $\texttt{p-send(m)}$ message to the $Sender$; this communication is denoted as $\texttt{\underline{in(m)}}$, where underline actions represent internal actions (similar to the $\tau$ of CCS), that cannot be further synchronised with the environment. At the other end process $Q$ receives the message from the $Receiver$. The holes $P$ and $Q$ can also have other interactions with their environment, represented here by actions $\texttt{p-a}$ and $\texttt{q-b}$. The underlying network is modelled by a medium entity transporting messages from the sender to the receiver, and that is able to detect transport errors, and signal them to the sender. The return $ack$ message from $receiver$ to $sender$ is supposed to be safe. The final transmission of the message to the recipient (the hole Q) includes the value of the ``error counter'' $ec$.

Fig. \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet of the system \emph{specification}:
the full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium.  
The pNet has a tree-like structure.  The root node of the tree {\it SimpleSystemSpec} is the top level of the pNet structure. It acts as the parallel operator. It consists of tree nodes: two Holes $P$ and $Q$ and one subnet, noted {\it PrefectBuffer}. Nodes of the tree are synchronized using four synchronization vectors, that  express the possible synchronization between the parameterized actions of a subset of the nodes. For instance, in the vector "$< \texttt{p-send(m)},\texttt{in(m)},\_> \rightarrow \texttt{\underline{in(m)}}$ only $P$ and {\it PrefectBuffer} nodes are involved in the synchronization. More precisely, the synchronization between these processes occurs when  process $P$  performs $\texttt{p-send(m)}$  action sending a message, and  the medium accepts the message through $\texttt{in(m)}$ action  at the same time. And the result that will be returned at upper level is the internal action  $\texttt{\underline{in(m)}}$.

@@

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition. The weak open transition shown here is the transition delivering the result of the algorithm to hole $Q$ by applying rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds an $\texttt{WT}_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {b1 \OTWeakarrow \tau b1}$; $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {b1 \OTWeakarrow \tau b1}$;  $\openrule
         {\{\}, True,
			() }
         {b1 \OTWeakarrow \tau b1}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+2) }
         {b1 \OTWeakarrow \tau b1}$. We can  can iterate this construction an arbitrary number of times, getting for any natural number $n$ the weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {\text{b}1 \OTWeakarrow \tau b1} \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{q-recv(ec)}  as $\alpha$, we get:
      $\openrule
         {\{\texttt{Q}\mapsto\texttt{q-revc(ec)}\subst{ec\gets ec+n}\}, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {b1 \OTWeakarrow {\underline{\texttt{out(m,ec)}}\subst{ec\gets ec+n}} b0} \forall n \geq 0$. 
\end{example}

  \begin{example}
    Figures \ref{SimpleProtCounter:WeakSpecOA} and \ref{SimpleProtCounter:ImplOA} respectively show the weak automata of the simple protocol specification and implementation.
    \TODO{comment...}

    For readability, we did not include the full details of the Implementation weak open transitions in Figure \ref{SimpleProtCounter:ImplOA}, they will appear below. In this automaton, many WOT are similar, and numbered accordingly as 3, 3', 3'', and 8, 8', 8'' respectively: they only differ by the name of the state variables in their respective source or target states.
\end{example}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=9cm]{XFIG/SimpleProtImpl-WOA}}
  \caption{Weak Open Automaton of the implementation}
   \label{SimpleProtCounter:ImplOA}
\end{figure}

@@

$post_3 = (\texttt{s1\_ec}\gets 0, \texttt{s1\_m}\gets \texttt{m})$

$post_4 = (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec})$

$post_5 = (\texttt{m2\_ec}\gets \texttt{m1\_ec})$

$post_6 = (\texttt{s1\_ec}\gets \texttt{m2\_ec+1})$

$post_7 = (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_ec}\gets \texttt{m1\_ec})$

@@

$Post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (\texttt{s1\_ec}\gets \texttt{s1\_ec}+1)$.
\medskip

So if we denote ${Post_{456}}^*$ any iteration of this loop, we get ${Post_{456}}^* = (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n)$ for any $n\ge 0$, and we will get e.g. the Post of the weak OT number 3' as $ post_4\shortotimes {Post_{456}}^*\shortotimes\ post_3 = (\texttt{m1\_msg}\gets m, \texttt{m1\_ec}\gets n), \forall n\ge 0$
\medskip

We can now show some of the weak OTs of Figure \ref{SimpleProtCounter:SpecOA-partial} (the full table is included in Appendix \ref{Appendix:FullExample}) :
\smallskip

$ 1 = \openrule
{\{\texttt{P}\mapsto \texttt{p-a}\}, [\forall \texttt{m}. \texttt{p-a} \neq \texttt{p-send(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ 3' = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{m1\_msg}\gets m, \texttt{m1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 210}
  \ \forall n\ge 0$

         $ 4 = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

$ 8 = \openrule
         {\{\}, True, ()}
         {201 \OTWeakarrow {\underline{\texttt{out(m,c)}}\subst{(m\gets \texttt{r1\_msg}, c\gets \texttt{r1\_ec}}} 202}
         $

\subsection{Composition properties: composition of weak open transitions}
We now have two different semantics for open pNets: a strong semantics, defined  as an open automaton, and as a weak semantics, defined as a weak open automaton. Like the open automaton, the weak open automaton features valuable composition properties. We can exhibit  a composition property and a decomposition property that relate open pNet composition with their semantics, defined as weak open automata. These are however technically more complex than the ones for open automata because each hole performs a set of actions, and thus a composed transition is the composition of one transition of the top-level pNet and a sequence of transitions of the sub-pNet that fills its hole. They can be found as Lemmas~\ref{lem-decomposeWOT}, Lemma~\ref{lem-Weakcompose1}, and Lemma~\ref{lem-Weakcompose} in~\ref{sec:app-composition}.

@@

\begin{lemma}[Alternative definition of weak open transitions]
\label{lem-rel-OT-WOT} Let $A = <J,\mathcal{S}, s_0,
    \mathcal{T}>$ be open automata and $<J,\mathcal{S}, s_0,
    \WT>$ be the
weak open automata derived from $A$.  The two following statements are equivalent
%For any weak open transition $\symb{WOT} \in \WT$, there exists an open transitions $\symb{OT} \in \mathcal{T}$ such that:\\
\begin{enumerate}
\item Either $   
\alpha = \tau \wedge  \set{\gamma}=\emptyset \wedge \Pred =\True \wedge \Post =\Id(s)\wedge s=s'$; or \\ 
there exist   $\set{\beta_{1i}}$, $\set{\beta_{2i}}$, and $\set{\beta_{3i}}$, $\Pred_{1i}$, $\Pred_{3i}$, $\Post_{1i}$, and  $\Pred_2$, $\Post_2$ s.t.:
\begin{mathpar}
%\mprset {vskip=.5ex}
%\inferrule 
\forall i \in [0..n].\openrule
    {
       \set{\beta_{1i}},\Pred_{1i},\Post_{1i}   }
         {s_{1i} \OTarrow {\tau} s_{1(i+1)}} \in \mathcal{T} \quad \wedge
\quad
\openrule
         {
           \set{\beta_2},\Pred_2,\Post_2 }
         {s_2 \OTarrow {\alpha} s'_2} \in \mathcal{T}
\quad \wedge\\
\forall i \in [0..m].\openrule
         {
           \set{\beta_{3i}},\Pred_{3i},\Post_{3i}    }
         {s_{3i} \OTarrow {\tau} s_{3(i+1)}} \in \mathcal{T}
 \end{mathpar}
\item  there exist $\set{\gamma}$, $\Pred$, $\Post$ s.t.
 \begin{mathpar}
%   \Longleftrightarrow  {
\openrule
         {
           \set{\gamma},
		\Pred, \Post
				 } {s \OTWeakarrow {\alpha'} s'} \in\WT
%}
\end{mathpar}
where\\
$
\alpha'=\alpha\displaystyle{\subst{\bigotimes_{j=n}^{0}\Post_{1j}}}\\
s=s_{10} \wedge s_{1(n+1)}=s_2 \wedge s'_2 = s_{30} \wedge s_{3(m+1)}=s'\\
\displaystyle{
\set{\gamma}=\mybigdotcup_{i=0}^n \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{0} \Post_{1j} } }  \dotcup  \vis{\set{\beta_2}\subst{\bigotimes_{j=n}^{0}\Post_{1j}}} \dotcup}\\
~\qquad \displaystyle{\mybigdotcup_{i=0}^m \vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{0}\Post_{1j}} }}
\\
\Pred=\bigwedge_{i=0}^n\Pred_{1i}\subst{\bigotimes_{j=i-1}^{0}\Post_{1j}}\land\Pred_2 \subst{\bigotimes_{j=n}^{0}\Post_{1j}}\land\\ 
~\qquad\quad\Big(\bigwedge_{i=0}^m\Pred_{3i}\subst{\bigotimes_{j=i-1}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{0}\Post_{1j}}\Big)\\
\Post=\bigotimes_{j=m}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{0}\Post_{1j}
$

@@

\begin{proof} ($\Rightarrow$) We proceed by induction on  $n$ and $m$ (where between each step either $m$ or $n$ increases).
\begin{itemize}
\item The base case there is one transition, so $n=0$ and $m=0$, we have:
\begin{mathpar}
 % \mprset {vskip=.5ex}
%\inferrule{
 \openrule
         {
           \set{\beta},\Pred,\Post}
         {s \OTarrow {\alpha} s'} \in \mathcal{T}
%}
%}
\end{mathpar} 
by rule \WTDeux we can directly conclude the implication: 
 \begin{mathpar}
 % \mprset {vskip=.5ex}
%\inferrule{
% \openrule
%         {
%           \set{\beta},\Pred,\Post}
%         {s \OTarrow {\alpha} s'} \in \mathcal{T}
%}

\openrule
         {
           \set{\beta},\Pred,\Post}
         {s \OTarrow {\alpha} s'} \in \mathcal{T} \Rightarrow 
{ \openrule
         {
           \vis{\set{\beta}}\!,\Pred,\Post
				 } {s \OTWeakarrow {\alpha} s'} \in \WT
}
\end{mathpar} 
\item For the inductive step, first we have by induction hypothesis that the formula holds for some lengths $m$ and $n$. Induction step is to infer that formula holds for transitions of length $m+1$ or $n+1$. 
We consider the case $n'=n+1$ (because $m'=m+1$ is similar).  We want to prove (1) $\Rightarrow$ (2) in the lemma, and in (1) we  focus on the case where there is a set of open transitions (this is the case: $s\neq s'$). In other words, we consider the sequence of ($n+m+2$) open transitions: 
\begin{mathpar}
\Big(\forall i \in [0..n+1].\openrule
    {
       \set{\beta_{1i}},\Pred_{1i},\Post_{1i}   }
         {s_{1i} \OTarrow {\tau} s_{1(i+1)}} \in \mathcal{T} \quad \wedge
\quad
\openrule
         {
           \set{\beta_2},\Pred_2,\Post_2 }
         {s_2 \OTarrow {\alpha} s'_2} \in \mathcal{T}
\quad \wedge\\
~~\qquad\qquad\forall i \in [0..m].\openrule
         {
           \set{\beta_{3i}},\Pred_{3i},\Post_{3i}   }
         {s_{3i} \OTarrow {\tau} s_{3(i+1)}} \in \mathcal{T}
\Big) 
\end{mathpar}

By recurrence hypothesis we suppose that the (1) $\Rightarrow$ (2) holds for $n$ and $m$ (compared to the line above, we remove the first $\tau$ transition). We have:\\

>>>>

\TODO{voisr si on redefinit les bigotimes pour prendre des maps et retourner des maps au lieu des substitutions (et on englobe dans une subst a la fin)}

\TODO{prendre soin des posts sur les actions partout}

===> dealine 14 Juillet

TODOs du 19 decembre

- enlever 5.2 et citer avocs

- ajouter des proof sketch dans le papier

- def 11 point 2:
A, Pred, Post
-----------   in T
s->s'
et A(j)=tau
implies blablabla

OLD OLD OLD (check)
Notes de reunion 24 Sept

- etat de l art: il faudrait des refs sur bisim en gemneral, probablement avec pas mal de pi calcul (Eric?), des refs sur pourquoi pnets et des refs sur utilite de la weak bisim en verif

- target: JLamp

Conventions
[14/9/2018]

Dans le texte pas de $\backslash$ devant pNet et pLTS

pNet devient P ou Q dans les equations; on garde $pLTS$

TODOS et deadline

Global and questions:
\begin{itemize}
\item \TODO{should we use iff/implies or LeftRightArrow/Rightarrow}
\item related works
\end{itemize}
Ludo:
\begin{itemize}
\item[X] reorga sec weak bisim + thm -> 10 Juil
\item[X]  intro et orga globale du papier -> fin aout
\item[partial] explication dans semantique open pnets et composition
\item uniformisation pNet P Q
\end{itemize}

@@

Fig. \ref{SimpleProt:Spec} (left) shows the example principle: two unspecified
processes $P$ and $Q$ (holes) communicate messages, with a data value
argument, through the two protocol entities. Process $P$ send an $\texttt{p\_send(m)}$ message to the $Sender$; this communication is denoted as $\texttt{\underline{in(m)}}$, where underline actions represent internal actions (similar to the $\tau$ of CCS), that cannot be further synchronised with the environment. At the other end process $Q$ receives the message from the $Receiver$. The holes $P$ and $Q$ can also have other interactions with their environment, represented here by actions $\texttt{p\_a}$ and $\texttt{q\_b}$. The underlying network is modelled by a medium entity transporting messages from the sender to the receiver, and that is able to detect transport errors, and signal them to the sender. The return $ack$ message from $receiver$ to $sender$ is supposed to be safe. The final transmission of the message to the recipient (the hole Q) includes the value of the ``error counter'' $ec$.

Fig. \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet of the system \emph{specification}:
the full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium.  
The pNet has a tree-like structure.  The root node of the tree {\it SimpleSystemSpec} is the top level of the pNet structure. It acts as the parallel operator. It consists of tree nodes: two Holes $P$ and $Q$ and one subnet, noted {\it PrefectBuffer}. Nodes of the tree are synchronized using four synchronization vectors, that  express the possible synchronization between the parameterized actions of a subset of the nodes. For instance, in the vector "$< \texttt{p\_send(m)},\texttt{in(m)},\_> \rightarrow \texttt{\underline{in(m)}}$ only $P$ and {\it PrefectBuffer} nodes are involved in the synchronization. More precisely, the synchronization between these processes occurs when  process $P$  performs $\texttt{p\_send(m)}$  action sending a message, and  the medium accepts the message through $\texttt{in(m)}$ action  at the same time. And the result that will be returned at upper level is the internal action  $\texttt{\underline{in(m)}}$.

@@

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition. The weak open transition here is the transition delivering the result of the algorithm to hole $Q$ by applying or rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds an $\texttt{WT}_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {b1 \OTWeakarrow \tau b1}$; $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {b1 \OTWeakarrow \tau b1}$;  $\openrule
         {\{\}, True,
			() }
         {b1 \OTWeakarrow \tau b1}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+2) }
         {b1 \OTWeakarrow \tau b1}$. We can  can iterate this construction an arbitrary number of times, getting for any natural number $n$ the weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {\text{b}1 \OTWeakarrow \tau b1} \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{q-recv(ec)}  as $\alpha$, we get:
      $\openrule
         {\{\texttt{Q}\mapsto\texttt{q-revc(ec)}\subst{ec\gets ec+n}\}, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {b1 \OTWeakarrow {\underline{\texttt{out(m,ec)}}\subst{ec\gets ec+n}} b0} \forall n \geq 0$. 
\end{example}

  \begin{example}
    Figures \ref{SimpleProtCounter:WeakSpecOA} and \ref{SimpleProtCounter:SpecOA-partial} respectively show the weak automata of the simple protocol specification and implementation.
    \TODO{comment...}
\end{example}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=9cm]{XFIG/SimpleProtImpl-WOA}}
  \caption{Weak Open Automaton of the implementation}
   \label{SimpleProtCounter:SpecOA-partial}
\end{figure}

%% \TODO{Eric: insert here a table with the details of the WOTs.Il manque encore au moins 3 WOT dans ce dessin...}

%% \smallskip
%% $1^st$ try, user-friendly style:

%% $$ 3 = \tau^*; \underline{in(m)}\{s1\_c:=0, s1\_m:=m\}; \tau\{s1\_c++\}^*$$

%% \smallskip
%% $2^nd$ try, full formal style ?:
%% \medskip

@@

$post_3 = (\texttt{s1\_c}\gets 0, \texttt{s1\_m}\gets \texttt{m})$

$post_4 = (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_c}\gets \texttt{s1\_c})$

$post_5 = (\texttt{m2\_c}\gets \texttt{m1\_c})$

$post_6 = (\texttt{s1\_c}\gets \texttt{m2\_c+1})$

$post_7 = (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m1\_c})$

@@

$Post_{456} = post_6 \shortotimes\ post_5 \shortotimes\ post_4
= (\texttt{s1\_c}\gets \texttt{s1\_c}+1)$.
\medskip

So if we denote ${Post_{456}}^*$ any iteration of this loop, we get ${Post_{456}}^* = (\texttt{s1\_c}\gets \texttt{s1\_c}+n)$ for any $n\ge 0$, and we will get e.g. the Post of the weak OT number 3' as $ post_4\shortotimes {Post_{456}}^*\shortotimes\ post_3 = (\texttt{m1\_msg}\gets m, \texttt{m1\_c}\gets n), \forall n\ge 0$
\medskip

We can now show some of the weak OTs of Figure \ref{SimpleProtCounter:SpecOA-partial} (the full table will be included in Appendix \ref{Appendix:FullExample}) :
\smallskip

$ 1 = \openrule
{\{\texttt{P}\mapsto \texttt{p\_a}\}, [\forall \texttt{m}. \texttt{p\_a} \neq \texttt{p\_send(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ 2 = \openrule
{\{\texttt{Q}\mapsto \texttt{q\_b}\}, [\forall \texttt{m}. \texttt{q\_b} \neq \texttt{q\_recv(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ 3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True,
    (\texttt{s1\_msg}\gets m, \texttt{s1\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  \ \forall n\ge 0$

$ 3' = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True,
    (\texttt{m1\_msg}\gets m, \texttt{m1\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 210}
  \ \forall n\ge 0$

$ 3'' = \openrule
  {\{\texttt{P}\mapsto \texttt{p\_send(m)}\}, True, 
    (\texttt{m2\_c}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 220}
  \ \forall n\ge 0$

         $ 4 = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_c}\gets \texttt{s1\_c}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

$ 5 = \openrule
         {\{\}, True, (\texttt{m2\_c}\gets \texttt{m1\_c}+n)}
         {210 \OTWeakarrow {\tau} 220}
         \ \forall n\ge 0$

$ 6 = \openrule
         {\{\}, True, (\texttt{s1\_c}\gets \texttt{m2\_c}+n)}
         {220 \OTWeakarrow {\tau} 100}
         \ \forall n\ge 0$

$ 7 = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_c}\gets \texttt{s1\_c}+n)}
         {100 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 0$

$ 7' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m1\_c}+n)}
         {210 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 0$

$ 7'' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_c}\gets \texttt{m2\_c}+n)}
         {220 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 1$

$ 8 = \openrule
         {\{\}, True, ()}
         {201 \OTWeakarrow {\underline{\texttt{out(m,c)}}\subst{(m\gets \texttt{r1\_msg}, c\gets \texttt{r1\_c}}} 202}
         $

\newpage
\subsection{Composition properties: composition of weak open transitions}
We now have two different semantics for open pNets: a strong semantics, defined  as an open automaton, and as a weak semantics, defined as a weak open automaton. Like the open automaton, the weak open automaton features valuable composition properties. We can exhibit  a composition property and a decomposition property that relate open pNet composition with their semantics, defined as weak open automata. These are however technically more complex than the ones for open automata because each hole performs a set of actions, and thus a composed transition is the composition of one transition of the top-level pNet and a sequence of transitions of the sub-pNet that fills its hole. They can be found as Lemmas~\ref{lem-decomposeWOT}, Lemma~\ref{lem-Weakcompose1}, and Lemma~\ref{lem-Weakcompose} in~\ref{sec:app-composition}.

@@

\begin{lemma}[Alternative definition of weak open transitions]
\label{lem-rel-OT-WOT} Let $A = <J,\mathcal{S}, s_0,
    \mathcal{T}>$ be open automata and $<J,\mathcal{S}, s_0,
    \WT>$ be the
weak open automata derived from $A$.  The two following statements are equivalent
%For any weak open transition $\symb{WOT} \in \WT$, there exists an open transitions $\symb{OT} \in \mathcal{T}$ such that:\\
\begin{enumerate}
\item Either $   
\alpha = \tau \wedge  \set{\gamma}=\emptyset \wedge \Pred =\True \wedge \Post =\Id(s)\wedge s=s'$; or \\ 
there exist   $\set{\beta_{1i}}$, $\set{\beta_{2i}}$, and $\set{\beta_{3i}}$, $\Pred_{1i}$, $\Pred_{3i}$, $\Post_{1i}$, and  $\Pred_2$, $\Post_2$ s.t.\footnote{$n=0$ (resp. $m=0$) corresponds to the case where there is no $\tau$ transition before (resp. after) the transition $\alpha$.}:
\begin{mathpar}
%\mprset {vskip=.5ex}
%\inferrule 
\forall i \in [1..n].\openrule
    {
       \set{\beta_{1i}},\Pred_{1i},\Post_{1i}   }
         {s_{1i} \OTarrow {\tau} s_{1(i+1)}} \in \mathcal{T} \quad \wedge
\quad
\openrule
         {
           \set{\beta_2},\Pred_2,\Post_2 }
         {s_2 \OTarrow {\alpha} s'_2} \in \mathcal{T}
\quad \wedge\\
\forall i \in [1..m].\openrule
         {
           \set{\beta_{3i}},\Pred_{3i},\Post_{3i}    }
         {s_{3i} \OTarrow {\tau} s_{3(i+1)}} \in \mathcal{T}
 \end{mathpar}
\item  there exist $\set{\gamma}$, $\Pred$, $\Post$ s.t.
 \begin{mathpar}
%   \Longleftrightarrow  {
\openrule
         {
           \set{\gamma},
		\Pred, \Post
				 } {s \OTWeakarrow {\alpha'} s'} \in\WT
%}
\end{mathpar}
where\\
$
\alpha'=\alpha\displaystyle{\subst{\bigotimes_{j=n}^{1}\Post_{1j}}}\\
s=s_{11} \wedge s_{1(n+1)}=s_2 \wedge s'_2 = s_{31} \wedge s_{3(m+1)}=s'\\
\displaystyle{
\set{\gamma}=\mybigdotcup_{i=1}^n \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{1} \Post_{1j} } }  \dotcup  \vis{\set{\beta_2}\subst{\bigotimes_{j=n}^{1}\Post_{1j}}} \dotcup}\\
~\qquad \displaystyle{\mybigdotcup_{i=1}^m \vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{1}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{1}\Post_{1j}} }}
\\
\Pred=\bigwedge_{i=1}^n\Pred_{1i}\subst{\bigotimes_{j=i-1}^{1}\Post_{1j}}\land\Pred_2 \subst{\bigotimes_{j=n}^{1}\Post_{1j}}\land\\ 
~\qquad\quad\Big(\bigwedge_{i=1}^m\Pred_{3i}\subst{\bigotimes_{j=i-1}^{1}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{1}\Post_{1j}}\Big)\\
\Post=\bigotimes_{j=m}^{1}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{1}\Post_{1j}
$

@@

\begin{proof} ($\Rightarrow$) We proceed by induction on  $n$ and $m$ (where between each step either $m$ or $n$ increases).
\begin{itemize}
\item In the base case there is one transition, which corresponds to $n=0$ and $m=0$, we have:
\begin{mathpar}
 % \mprset {vskip=.5ex}
%\inferrule{
 \openrule
         {
           \set{\beta},\Pred,\Post}
         {s \OTarrow {\alpha} s'} \in \mathcal{T}
%}
%}
\end{mathpar} 
by rule \WTDeux\ we can directly conclude the implication: 
 \begin{mathpar}
 % \mprset {vskip=.5ex}
%\inferrule{
% \openrule
%         {
%           \set{\beta},\Pred,\Post}
%         {s \OTarrow {\alpha} s'} \in \mathcal{T}
%}

\openrule
         {
           \set{\beta},\Pred,\Post}
         {s \OTarrow {\alpha} s'} \in \mathcal{T} \Rightarrow 
{ \openrule
         {
           \vis{\set{\beta}}\!,\Pred,\Post
				 } {s \OTWeakarrow {\alpha} s'} \in \WT
}
\end{mathpar} 
\item For the inductive step, first we have by induction hypothesis that the formula holds for some lengths $m$ and $n$. Induction step is to infer that formula holds for transitions of length $m+1$ or $n+1$. 
We consider the case $n'=n+1$ ($m'=m+1$ is similar).  We want to prove (1) $\Rightarrow$ (2) in the lemma, and in (1) we  focus on the case where there is a set of open transitions (this is the case: $s\neq s'$). In other words, we consider the sequence of ($n+m+2$) open transitions: 
\begin{mathpar}
\Big(\forall i \in [1..n+1].\openrule
    {
       \set{\beta_{1i}},\Pred_{1i},\Post_{1i}   }
         {s_{1i} \OTarrow {\tau} s_{1(i+1)}} \in \mathcal{T} \quad \wedge
\quad
\openrule
         {
           \set{\beta_2},\Pred_2,\Post_2 }
         {s_2 \OTarrow {\alpha} s'_2} \in \mathcal{T}
\quad \wedge\\
~~\qquad\qquad\forall i \in [1..m].\openrule
         {
           \set{\beta_{3i}},\Pred_{3i},\Post_{3i}   }
         {s_{3i} \OTarrow {\tau} s_{3(i+1)}} \in \mathcal{T}
\Big) 
\end{mathpar}

By recurrence hypothesis we suppose that  (1) $\Rightarrow$ (2) holds for $n$ and $m$ (compared to the line above, we remove the first $\tau$ transition). We have:\\

$
%\mprset {vskip=.5ex}
%\inferrule 
\Big(\forall i \in [2..n+1].\openrule
    {
       \set{\beta_{1i}},\Pred_{1i},\Post_{1i}   }
         {s_{1i} \OTarrow {\tau} s_{1(i+1)}} \in \mathcal{T} \,\wedge\,
\openrule
         {
           \set{\beta_2},\Pred_2,\Post_2 }
         {s_2 \OTarrow {\alpha} s'_2} \in \mathcal{T}
\, \wedge\\
{}~~\hspace{2.25cm}\forall i \in [1..m].\openrule
         {
           \set{\beta_{3i}},\Pred_{3i},\Post_{3i}   }
         {s_{3i} \OTarrow {\tau} s_{3(i+1)}}\! \in\! \mathcal{T}
\Big)  \Rightarrow\!  {
\openrule
         {
           \set{\gamma},
		\Pred, \Post
				 } {s'' \OTWeakarrow {\alpha'} s'} \in\WT
}
$
where\\
$
\alpha'=\alpha\displaystyle{\subst{\bigotimes_{j=n+1}^{2}\!\!\Post_{1j}}}\\
s''=s_{12} \wedge s_{1(n+2)}=s_2 \wedge s'_2 = s_{31} \wedge s_{3(m+1)}=s'\\
\displaystyle{
\set{\gamma}=\mybigdotcup_{i=2}^{n+1} \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{1} \Post_{1j} } }  \dotcup  \vis{\set{\beta_2}\subst{\bigotimes_{j=n+1}^{2}\Post_{1j}}} \dotcup}
\\
{}~\qquad 
\displaystyle{\mybigdotcup_{i=1}^m \vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{1}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{2}\Post_{1j}} }}
\\
\Pred=\bigwedge_{i=2}^{n+1}\Pred_{1i}\subst{\bigotimes_{j=i-1}^{1}\Post_{1j}}\land\Pred_2 \subst{\bigotimes_{j={n+1}}^{2}\Post_{1j}}\land
\\ 
{}~\qquad\quad
\Big(\bigwedge_{i=1}^m\Pred_{3i}\subst{\bigotimes_{j=i-1}^{1}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j={n+1}}^{2}\Post_{1j}}\Big)\\
\Post=\bigotimes_{j=m}^{1}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j={n+1}}^{2}\Post_{1j}
$
